A Methodology and Implementation of Automated Emissions Harmonization for Use in Integrated Assessment Models

### 2017-06-07 elsfrontmatter.tex
Shinichiro_edition_7-June-2017
Update on Overleaf.

\author[iiasa,nies]{Shinichiro Fujimori}

@@

\begin{abstract}
Harmonization describes the process of calibrating Integrated Assessment Model
(IAM) GHG and air pollutants results with a new source of historical emissions trajectories. To date,
harmonization has been performed separately by individual modeling teams through
% * <sfujimori112256@gmail.com> 2017-05-24T19:26:26.142Z:
% 
% > harmonization has been performed separately by individual modeling teams 
% I think harmonization is not done. We just report the direct model outputs.
% 
% One of the ways to  make the background is that earlier IPCC scenarios (SRES and RCPs) did not document the harmonization method and far from transparent.
% 
% ^.
the use of expert opinion and solicitation. As models and their results become
more complex in both regional and sectoral dimensions, an automated approach to
choosing harmonization methods becomes necessary. This work describes a novel
% * <sfujimori112256@gmail.com> 2017-05-24T20:13:08.493Z:
% 
% > novel
% It would be better to stress how novel and useful.
% 
% ^.
methodology for determining such harmonization methods and an open-source Python
code base for implementing the methodology. Results are shown for two example
scenarios using the MESSAGE-GLOBIOM IAM that satisfactorily harmonize over 98\%
of the total emissions trajectories.
\end{abstract}

>>>>

\author[iiasa,nies]{Shinichiro Fujimori}
\author[iiasa]{Keywan Riahi}

@@

\begin{abstract}
Harmonization describes the process of calibrating Integrated Assessment Model
(IAM) GHG and air pollutants results with a new source of historical emissions trajectories. To date,
harmonization has been performed separately by individual modeling teams through
% * <sfujimori112256@gmail.com> 2017-05-24T19:26:26.142Z:
% 
% > harmonization has been performed separately by individual modeling teams 
% I think harmonization is not done. We just report the direct model outputs.
% 
% One of the ways to  make the background is that earlier IPCC scenarios (SRES and RCPs) did not document the harmonization method and far from transparent.
% 
% ^.
the use of expert opinion and solicitation. As models and their results become
more complex in both regional and sectoral dimensions, an automated approach to
choosing harmonization methods becomes necessary. This work describes a novel
% * <sfujimori112256@gmail.com> 2017-05-24T20:13:08.493Z:
% 
% > novel
% It would be better to stress how novel and useful.
% 
% ^.
methodology for determining such harmonization methods and an open-source Python
code base for implementing the methodology. Results are shown for two example
scenarios (with and without climate policy cases) using the MESSAGE-GLOBIOM IAM that satisfactorily harmonize over 98\%
of the total emissions trajectories.
\end{abstract}

### 2017-06-07 intro.tex
Update on Overleaf.

Integrated Assessment Models (IAMs) are tools used to understand the complex
interactions between energy systems, economic systems, land use, the climate,
and air pollution, among other sectors . IAMs provide global projections of
systemic change by dividing the world into a number of representative regions
(typically 10 to 30), the definition of which is distinct for each model
\cite{krey_global_2014}. Results from IAMs are integral in a number of
international studies, which notably include projections of climate and economic
futures. Recently, flagship IAMs have developed scenarios based on the Shared
Socioeconomic Pathways (SSPs) \cite{van_vuuren_energy_2017, fricko_marker_2017,
  fujimori_ssp3:_2017, calvin_ssp4:_2017, kriegler_fossil-fueled_2017} which
quantify a variety of potential global futures. The SSPs are primarily designed
to be used in analyses of broad climate research which include earth system
model (ESM) simulations, climate impact, adaptation climate mitigation
studies. They consequently underpin the United Nations' Intergovernmental Panel
on Climate Change (IPCC) across working groups.

@@

With some frequency, however, historical data is updated to incorporate new
assumptions of source activity, new emissions accounting practices, or other
concerns. The global climate change community has recently developed a standard
historical emissions dataset for both anthropogenic emissions (i.e., the
Community Emissions Datasystem (CEDS)) \cite{hoesly_historical_2017} and
land-use change (LUC) emissions \cite{van_marle_historic_2017} which, in
conjunction with the SSP IAM trajectories, will be used for climate-related
modeling exercises of CMIP6.

>>>>

Integrated Assessment Models (IAMs) are tools used to understand the complex
interactions between energy systems, economic systems, land use, the climate,
and air pollution, among other sectors . IAMs provide global projections of
systemic change by dividing the world into a number of representative regions
(typically 10 to 30), the definition of which is distinct for each model
\cite{krey_global_2014}. Results from IAMs are integral in a number of
international studies, which notably include projections of climate and economic
futures. Recently, flagship IAMs have developed scenarios based on the Shared
Socioeconomic Pathways (SSPs) \cite{van_vuuren_energy_2017, fricko_marker_2017,
  fujimori_ssp3:_2017, calvin_ssp4:_2017, kriegler_fossil-fueled_2017} which
quantify a variety of potential global futures. The SSPs are primarily designed
to be used in analyses of broad climate research which include earth system
model (ESM) simulations, climate impact, adaptation and climate mitigation
studies. They consequently underpin the United Nations' Intergovernmental Panel
on Climate Change (IPCC) across working groups.

@@

With some frequency, however, historical data is updated to incorporate new
assumptions of source activity, new emissions accounting practices, or other
concerns. The global climate change community has recently developed a standard
historical emissions dataset for both anthropogenic emissions (i.e., the
Community Emissions Datasystem (CEDS) \cite{hoesly_historical_2017} and
land-use change (LUC) emissions \cite{van_marle_historic_2017}) which, in
% * <sfujimori112256@gmail.com> 2017-06-07T09:59:35.820Z:
% 
% > land-use change (LUC) emissions \cite{van_marle_historic_2017}
% I think this one also anthropogenic
% 
% ^.
conjunction with the SSP IAM trajectories, will be used for climate-related
modeling exercises of CMIP6.

### 2017-06-07 meth.tex
Update on Overleaf.

A number of characteristics impact the decision of which default method to
select based on the effect of the characteristic on the potential harmonized
trajectory. For example, it is possible for models to report zero values in the
harmonization year in situations in which technologies are introduced in future
time periods in regions or for sectors which produce an emissions species that
is absent in the initial modeling period. In such cases, an offset method is
required as a ratio method would mask future emissions and erroneously harmonize
the trajectory. In most cases, however, models do report values in the
harmonization year. When model and historical values are relatively close, a
convergence method is chosen in order to be as representative as possible to the
underlying unharmonized model results. The final model-based characteristic of
import is whether the model reports negative emissions. Such a case is possible
for gas species which can be extracted from the environment and stored, as is
the case for CO2 in future scenarios with high climate mitigation. If a model
provides a trajectory that transitions from positive to negative emissions, then
a convergence method is used in order to guarantee capture of this
transition. Otherwise, it is possible, given the relative values between history
and model in the harmonization year, for a negative trajectory to be
inappropriately harmonized to a positive, but decreasing, trajectory.

@@

The full \code{aneris} workflow is comprised of a number of components shown
graphically in Figure \ref{fig:workflow}. Unharmonized model data and a
run-control configuration are read in via an Excel spreadsheet. Data is assumed
to be in the \textit{IAM Consortium} format, i.e., using \code{Model},
\code{Scenario}, \code{Region}, \code{Variable}, and \code{Unit} columns in
addition to columns representing each modeled time period. Users are able to
control the harmonization process via a number of options. First, users can
provide \textit{override} methods for any combination of region and variable
(i.e., sector and gas species). If override methods are provided, they are used
instead of the default methods as determined by the default method decision
tree. Furthermore, users can set the above-mentioned thresholds as well as the
LUC method used in the decision tree (see Figure
\ref{fig:decision_tree}). Further detail of input parameters can be found at
\hyperlink{http://aneris.readthedocs.io/en/latest/}{http://aneris.readthedocs.io/en/latest/}.

@@

Input data then undergoes a cleaning operation, which adds (null) model
trajectories that exist in the historical dataset but are not provided by the
model input and detects any issues that would cause the harmonization process to
fail. The methods used to harmonize the data are then determined and the
harmonization process is executed. Upon completion of the harmonization process,
aggregation of common analysis regions is performed (e.g., the 5 RCP
regions). Finally, any exogenous trajectories the user provides are
added. Exogenous trajectories are normally provided for unmodeled gases with
well-accepted scenario trajectories, e.g., chlorofluorocarbons provided by WMO
\cite{wmo2014}. Upon completion, the harmonized trajectories and meta data
regarding the harmonization process are returned. A description of all returned
meta data is provided in Table \ref{tab:metadata}.

>>>>

A number of characteristics impact the decision of which default method to
select based on the effect of the characteristic on the potential harmonized
trajectory. For example, it is possible for models to report zero values in the
harmonization year in situations in which technologies are introduced in future
time periods in regions or for sectors which produce an emissions species that
is absent in the initial modeling period. In such cases, an offset method is
required as a ratio method would mask future emissions and erroneously harmonize
the trajectory. In most cases, however, models do report values in the
harmonization year. When model and historical values are relatively close, a
convergence method is chosen in order to be as representative as possible to the
underlying unharmonized model results. The final model-based characteristic of
import is whether the model reports negative emissions. Such a case is possible
for gas species which can be extracted from the environment and stored, as is
the case for CO2 in future scenarios with climate mitigation. If a model
% * <sfujimori112256@gmail.com> 2017-06-07T10:07:21.265Z:
% 
% > with climate mitigation
% Not necessarily high mitigation. Land use can be negative even in baseline.
% 
% ^.
provides a trajectory that transitions from positive to negative emissions, then
a convergence method is used in order to guarantee capture of this
transition. Otherwise, it is possible, given the relative values between history
and model in the harmonization year, for a negative trajectory to be
inappropriately harmonized to a positive, but decreasing, trajectory.

@@

The full \code{aneris} workflow is comprised of a number of components shown
graphically in Figure \ref{fig:workflow}. Unharmonized model data and a
run-control configuration are read in via an Excel spreadsheet. Data is assumed
to be in the \textit{IAM Consortium} format, i.e., using \code{Model},
\code{Scenario}, \code{Region}, \code{Variable}, and \code{Unit} columns in
addition to columns representing each modeled time period. Users are able to
control the harmonization process via a number of options. First, users can
provide \textit{override} methods for any combination of region and variable
(i.e., sector and gas species). If override methods are provided, they are used
% * <sfujimori112256@gmail.com> 2017-06-07T10:12:45.308Z:
% 
% Maybe it would be better to add an explanation why we allow this option like
% "We allow this overriding option because the default method do not meet our criteria sometimes (but usually very few cases)"
% 
% ^.
instead of the default methods as determined by the default method decision
tree. Furthermore, users can set the above-mentioned thresholds as well as the
LUC method used in the decision tree (see Figure
\ref{fig:decision_tree}). Further detail of input parameters can be found at
\hyperlink{http://aneris.readthedocs.io/en/latest/}{http://aneris.readthedocs.io/en/latest/}.

@@

Input data then undergoes a cleaning operation, which adds (null) model
trajectories that exist in the historical dataset but are not provided by the
model input and detects any issues that would cause the harmonization process to
fail. The methods used to harmonize the data are then determined and the
harmonization process is executed. Upon completion of the harmonization process,
aggregation of common analysis regions is performed (e.g., the 5 RCP
regions). Finally, any exogenous trajectories the user provides are
% * <sfujimori112256@gmail.com> 2017-06-07T10:17:07.935Z:
% 
% > (e.g., the 5 RCP
% > regions) 
% It might be good to show what is 5 regions (or definition) in the SM.
% 
% ^.
added. Exogenous trajectories are normally provided for unmodeled gases with
well-accepted scenario trajectories, e.g., chlorofluorocarbons provided by WMO
\cite{wmo2014}. Upon completion, the harmonized trajectories and meta data
regarding the harmonization process are returned. A description of all returned
meta data is provided in Table \ref{tab:metadata}.
